---
title: "p8105_hw1_sd3731"
author: "Susie Dong"
date: "2023-09-23"
output: github_document
---
# Problem 1

```{r, message=FALSE}
## load the dataset
library(moderndive)
library(tidyverse)
data("early_january_weather")
head(early_january_weather, 5)
```

## describe the dataset

This is data frame of 358 rows representing hourly measurements and 15 variables.

### variables name

```{r}
colnames(early_january_weather)
```

### Row Size

```{r}
nrow(early_january_weather)
ncol(early_january_weather)
```

### Mean Temperature

```{r}
mean(pull(early_january_weather, temp))
```

## scatterplot

### create a scatterplot

```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point() + 
  ggtitle("Temperature - Time Hour Scatter Plot") + theme(plot.title = element_text(hjust = 0.5))

ggsave("hw1_scatterplot_p1.png")
```

### discribe a scatterplot

-The scatterplot demonstrates a distinct diurnal temperature variation, characterized by daytime temperature increases and nighttime decreases, aligning with typical daily temperature fluctuations.

-The color-coded points reveal variations in humidity levels with time and temperature, with darker colors indicating increased humidity at specific times, shedding light on humidity fluctuations.

-The existence of clustered points with similar colors hints at potential groupings and patterns, potentially signifying specific times of the day when humidity levels tend to be alike, prompting a deeper exploration of underlying weather or climate patterns.


# Problem 2

Create a data frame comprised of:

-a random sample of size 10 from a standard Normal distribution
-a logical vector indicating whether elements of the sample are greater than 0
-a character vector of length 10
-a factor vector of length 10, with 3 different factor “levels”
-Try to take the mean of each variable in your dataframe. What works and what doesn’t?

## create a data frame

```{r}
sample_data = rnorm(10)

df = tibble(
  vec_numeric = sample_data,
  vec_logic = sample_data > 0,
  vec_str = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(sample(c("A", "B", "C"), 10, replace = TRUE))
)
df
```

## the mean of each variable

```{r}
mean_numeric = mean(pull(df, vec_numeric))
mean_lv = mean(pull(df, vec_logic))
mean_cv = mean(pull(df, vec_str))
mean_fac = mean(pull(df, vec_factor))
```
When calculating the mean for the four variables, the numeric and logical vectors produced numeric results, whereas the character vector and the factor vector yielded "NA."

```{r}
mean(as.numeric(df$log_vec))
mean(as.numeric(df$cha_vec))
mean(as.numeric(df$fac_vec))
```

## results

- The mean of numeric and logical variables can be successfully calculated, while the means of character and factor variables cannot be calculated (they return NA).

- After coercive conversion with as.numeric, the logical and factor variables become compatible to calculate the mean. However, for the character variable, it becomes compatible to calculate the mean only when every element of this variable consists solely of digits; otherwise, it returns NA.

## Explanation

-Calculating the mean is straightforward for numeric and logical variables. Numeric variables inherently consist of numbers that can be used in arithmetic calculations, while logical variables automatically convert to 0 or 1 based on their boolean values.

-By using the as.numeric function, a logical variable can be forcefully transformed into 0 or 1, and this conversion doesn't affect the mean calculation result.

-When the as.numeric function is applied to a factor variable, it is transformed into natural numbers (0, 1, 2, etc.) according to the lexicographic order of its levels, enabling it to be used in numeric calculations.

-In the case of character variables, the as.numeric function can coerce them into numeric values only if every element within the variable consists solely of digits. Otherwise, the conversion process returns NA, and any arithmetic calculation using these values also results in NA.
- the different levels in factor sequence becomes a sequence with a
start number 1.

## Conclusion

- When attempting arithmetic calculations, such as calculating the mean, the outcome varies significantly based on the variable type. However, by using the as.numeric function, it becomes possible to forcefully convert logical and factor variables into numeric values, making them suitable for arithmetic operations. In the case of character variables, only those containing digits can be converted to numeric values; otherwise, the conversion results in NA.
